---
title: 'class09: Unsupervised Learning Mini-Project'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Hands-on Worksheet##


#Section 1: Data Analysis

```{r}
wisc.df <- read.csv("WisconsinCancer_File.csv")
head(wisc.df)
```

We have `r nrow(wisc.df)` samples in this dataset

```{r}
nrow(wisc.df)
```


How many benign  and malignant samples do we have in the dataset?
```{r}
table(wisc.df$diagnosis)
```

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```


Store the diagnosis for reference in the future was a saparate vector

```{r}
diagnosis <- wisc.df$diagnosis
```


#Questions
Q1. How many observations are in this dataset?
```{r}
nrow(wisc.df)
```
    569
    
Q2. How many of the observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```
    212
    
Q3. How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.df)
```

```{r}
(grep("_mean", colnames(wisc.df), value = TRUE))
```

Number of variables we are looking for
```{r}
length(grep("_mean", colnames(wisc.df), value = TRUE))
```



## Section 2: Performing PCA
The next step in your analysis is to perform principal component analysis (PCA) on wisc.data.

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.
The input variables have significantly different variances.
Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before.
```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)

```

```{r}
round(colMeans(wisc.data), 3)
```

```{r}
round( apply(wisc.data, 2, sd), 3)
```

Values look very different so I will use `scale=TRUE` when I run PCA

```{r}
#Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale= TRUE)
summary(wisc.pr)
```

```{r}
plot(wisc.pr)
```

Let's make a plot of PC1 vs. PC2

```{r}
plot(wisc.pr$x[, 1], wisc.pr$x[, 2])
```


Color by benign/malignant...
```{r}
plot(wisc.pr$x[, 1], wisc.pr$x[, 2], col=diagnosis, xlab="PC1", ylab = "PC2")
```




Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

##Unsure code to determine Q4-Q6##
```{r}
x <- summary(wisc.pr)
x$importance[,1]
```

```{r}
which(x$importance[, 1] >0.9)
```
##_______________________________________________________________________##


### New Package for Plotting

**ggplot based graph**

install.packages("factoextra")
library(factoextra)

```{r}
fviz_eig(wisc.pr, addlabels = TRUE)
```


Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?


```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, "complete" )
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2")
```



```{r}
plot(wisc.pr.hclust)

```

#### Look at clustering of benign and malignant
```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```
```{r}
table(grps, diagnosis)
```
**28 people in group 1 are false positives; 24 people in group 2 are false negatives**

##### Section 7: Prediction

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

Plot these new samples on our PC1 vs PC2 plot

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)
text(npc[,1], npc[,2], labels = c(1,2), col="white")
```

Q17. Which of these new patients should we prioritize for follow up based on your result


### Installing rgl 3D cluster plot representation (Section 5) ###
```{r}
install.packages("rgl")
```


```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)
```

###__________________________________________________________________________________________________###
