---
title: "Class 11: Structural Bioinformatics Part 1 "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structure data



>Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

>Determine what proportion of structures are protein? 

>Q2: How many HIV-1 protease structures are in the current PDB?

```{r}
# Read CSV
info <- read.csv("Data Export Summary.csv")
View(info)

```

Total number of entries
```{r}
sum(info$Total)
```

Proportion of entries from each method
```{r}
(info$Total/sum(info$Total))*100
```

Proportion of structures that are protein
```{r}
round((sum(info$Proteins)/ sum(info$Total))*100, 2)
```

##HIV Protein Structure Analysis

Here we will read the 1HSG PDB structure and select the protein component and write out a new **protein-only** PDB formt file. We then do the same for the ligand (i.e. known drug molecule) creating a **ligand-only** PDB file.
```{r}
library(bio3d)

pdb <- read.pdb("1hsg.pdb")
pdb
#read.pdb()
#atom.select()
#write.pdb()
#trim.pdb()
```


```{r}
protein <- atom.select(pdb,"protein", value=TRUE)
write.pdb(protein, file="1hsg_protein.pdb")
```

```{r}
ligand <- atom.select(pdb,"ligand", value=TRUE)
write.pdb(ligand, file="1hsg_ligand.pdb")
```

