---
title: "class14_RMarkdown"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load DESeq2
```{r}
library(DESeq2)
```

#Load our data
```{r}
metaFile <- "GSE37704_metadata.csv"
countFile <- "GSE37704_featurecounts.csv"

# Import metadata and take a peak
colData = read.csv(metaFile, row.names=1)
head(colData)
```

```{r}
# Import countdata
countData = read.csv(countFile, row.names=1)

#We need the countData and colData files to match up so we will need to remove that odd first column in countData namely contData$length.

countData <- as.matrix(countData[,2:7])
countData = countData[rowSums(countData)!=0, ]
head(countData)
```

#Setup the DESeqDataSet object required for the DESeq() function and then run the DESeq pipeline. This is again similar to our last days hands-on session.
```{r}
dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)
dds

```

#Next, get results for the HoxA1 knockdown versus control siRNA (remember that these were labeled as "hoxa1_kd" and "control_sirna" in our original colData metaFile input to DESeq, you can check this above and by running resultsNames(dds) command).

```{r}
res = results(dds, contrast=c("condition", "hoxa1_kd", "control_sirna"))
summary(res)
```

#Now we will make a volcano plot, a commonly produced visualization from this type of data that we introduced last day. Basically it's a plot of log2 fold change vs -log adjusted p-value.

```{r}
plot( res$log2FoldChange, -log(res$padj) )

# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
inds <- (abs(-log(res$padj)) > 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange, -log(res$padj), col= mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.01), col="gray", lty=2)
```

#Since we mapped and counted against the Ensembl annotation, our results only have information about Ensembl gene IDs. However, our pathway analysis downstream will use KEGG pathways, and genes in KEGG pathways are annotated with Entrez gene IDs.
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)
```


